---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---

Reproducible Research: Peer Assessment 1
========================================

First we load all packages needed:
```{r, loadingPackages, echo=TRUE, results="asis"}
require("knitr")
require('xtable')
require("data.table")
require("ggplot2")
```

Next we assure that date format is in English. It is Linux
specific, so if you run it on Windows it could not work.
```{r, setlocale, echo=TRUE, results="asis" }
Sys.setlocale("LC_TIME", 'en_US.UTF-8')
```

## Loading and preprocessing the data

```{r, loadingData, echo=TRUE, results="asis"}
dataDir <- "data"
if (!file.exists(dataDir)) {dir.create(dataDir)}
unzip("activity.zip", exdir=dataDir)
activity.df <- read.csv(paste0(dataDir, "/activity.csv"))
activty.head <- xtable(head(activity.df))
print(activty.head, type="html")
```


## What is mean total number of steps taken per day?

### Head of the table with total number of steps per day (we have removed rows with NA).
```{r total_steps_per_day, echo=TRUE, results="asis"}
activity.dt <- as.data.table(activity.df)
steps <- activity.dt[!is.na(steps),.(number=sum(steps)),by=date]
print(xtable(head(steps)), type="html")
```

### Histogram of the total number of steps taken each day.
```{r histograma, echo=TRUE}
steps.mean <- mean(steps$number)
steps.median <- median(steps$number)

steps.gg1 <- ggplot(steps, aes(x=number))
steps.gg1+geom_histogram(binwidth = 850) + geom_vline(xintercept = steps.median, colour = "red")  
steps.gg2 <- ggplot(steps, aes(x=factor(0),y=number))
steps.gg2 + geom_boxplot()+coord_flip()
```

### Mean and median of the total number of steps taken per day.
```{r mean_median, echo=TRUE}
steps.mean
steps.median
```
Mean is `r sprintf("%.2f", steps.mean)` and median `r steps.median`.


## What is the average daily activity pattern?

### Time series plot

Plot of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all days (y-axis)

```{r}
steps2 <- activity.dt[,.(average=mean(steps, na.rm=TRUE),median=median(steps, na.rm=TRUE)),by=interval]

interval.max <- steps2[which.max(steps2[,average]),interval]
average.max <- steps2[which.max(steps2[,average]),average]

ggplot(steps2, aes(x=interval, y=average))+geom_line(colour="black")+
    geom_vline(xintercept=interval.max, colour="blue")+
        geom_hline(yintercept=average.max, colour="red")
```

### 5-minute interval that contains the maximum number of steps

`r interval.max`

## Imputing missing values

### Total number of missing values in the dataset
```{r}
sum(activity.dt[,is.na(steps)])
```

### Create a new dataset that is equal to the original dataset but with the missing data filled in.

We use **mean for 5-minute interval** for filling in all of the missing values in the dataset.


```{r}
#In order to clone we wrap by data.table
activity.nona <- data.table(activity.dt)

steps.average <- function(int) {
  steps2[interval == int,average]
}

activity.nona[,steps := ifelse(is.na(steps), steps.average(interval), steps)]
steps.total.per.day <- activity.dt[,.(total=sum(steps, na.rm=TRUE)), by=date]
steps.total.per.day.nona <- activity.nona[,.(total=sum(steps)), by=date]
```

### Make a histogram of the total number of steps taken each day and Calculate and report the mean and median total number of steps taken per day.

```{r}
ggplot(steps.total.per.day.nona)+geom_histogram(aes(x=total),binwidth=1000)
#qplot(total, data=steps.total.per.day, binwidth=1000)

steps.average.per.day <- activity.nona[,.(average=mean(steps)), by=date]
steps.median.per.day <- activity.nona[,.(median=median(steps)), by=date]
steps.mean.nona <- mean(steps.total.per.day.nona$total)
steps.median.nona <- median(steps.total.per.day.nona$total)
```



After filling in NA , mean is `r sprintf("%.2f",steps.mean.nona)`
and median is `r sprintf("%.2f",steps.median.nona)`.
Before mean wss `r sprintf("%.2f", steps.mean)` and median `r steps.median`.

### Impact of imputing missing data

```{r}
ggplot()+ geom_line(data=steps.total.per.day,aes(x=as.Date(date),y=steps.total.per.day.nona$total,color='blue'), alpha=0.4, size=1)+
    geom_line(data=steps.total.per.day,aes(x=as.Date(date),y=total,color='red'), alpha=0.4, size=0.8)+
    xlab('Date')+ylab('Steps per day') + 
    scale_colour_manual(name = '', 
         values =c('blue'='blue','red'='red'), labels = c('NA removed','with NA'))

```

## Are there differences in activity patterns between weekdays and weekends?


### Panel plot containing time series plots
```{r}
activity.nona[,day:=as.factor(
                   ifelse(
                       (weekdays(as.Date(date)) %in% c("Saturday","Sunday")),
                       "weekend","weekday"))
              ]

steps3 <- activity.nona[,.(average=mean(steps, na.rm=TRUE)),
                        by=.(interval, day)]
ggplot(steps3, aes(x=interval, y = average)) + geom_line()+facet_grid(day ~ .)
```
